---
title: "Extract polygons from dense cell segmentation"
author: "Shila Ghazanfar"
date: "15/06/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load functions

```{r}
source("../scripts/segmentation_functions.R")
```

# Load some example data

```{r}
data = readRDS("../data/dense_segmentation_example.Rds")
```

This is a dense cell segmentation per pixel, which also contains some sparse 
segments that do not correspond to real cells.

```{r}
dim(data)
data[1:5,1:5]
ncells = length(unique(c(data)))
ncells

col = sample(rainbow(ncells))
image(data, col = col)
```

# Extract polygons

This function will extract the smallest set of vertices that can accurately 
reflect the polygon structure of the segmentation. This method only works for
solid segments (no doughnuts!).

```{r}
boundary_polygons = extractPolygonsFromSegmentation(
    data,
    modes = c("boundaryPolygons"),
    cellFilterFactor = 5,
    minPixels = 50,
    cellPrefix = "cell_",
    cellSuffix = "",
    plot = TRUE,
    verbose = FALSE
)
colnames(boundary_polygons) <- c("x","y","cell","uniqueID")

dim(boundary_polygons)
head(boundary_polygons)

boundary_polygons <- pruneStraightLines(boundary_polygons, group = "uniqueID")

dim(boundary_polygons)
head(boundary_polygons)

g1 = ggplot(boundary_polygons, aes(x = x, y = y, group = uniqueID)) + 
    geom_polygon(aes(fill = uniqueID), alpha = 0.5, colour = "black", size = 0.3) + 
    theme_classic() + 
    theme(legend.position = "none") +
    ggtitle("Extracted polygons using dense segmentation") +
    NULL
g1
```

# Faster (but less accurate) approach using convex hull

```{r}
boundary_polygons_chull = extractPolygonsFromSegmentation(
    data,
    modes = c("chull"),
    cellFilterFactor = 5,
    minPixels = 50,
    cellPrefix = "cell_",
    cellSuffix = "",
    plot = TRUE,
    verbose = FALSE
)
# much faster but not as clean
colnames(boundary_polygons_chull) <- c("x","y","z", "value", "cell","uniqueID")

dim(boundary_polygons_chull)
head(boundary_polygons_chull)

g2 = ggplot(boundary_polygons_chull, aes(x = x, y = y, group = uniqueID)) + 
    geom_polygon(aes(fill = uniqueID), alpha = 0.5, colour = "black", size = 0.3) + 
    theme_classic() + 
    theme(legend.position = "none") +
    ggtitle("Extracted polygons using convex hull") +
    NULL
g2
```

# Side-by-side

```{r}
g1 + g2
```

```{r}
sessionInfo()
```