---
title: "Example of plotVectorField usage"
author: "Shila Ghazanfar"
date: "30/07/2020"
output:
       html_document:
                     toc: true
                     toc_float:
                           collapsed: false
                           smooth_scroll: false
                     code_folding: hide
                     fig_width: 10
                     fig_height: 8
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE,
                      cache = FALSE, cache.lazy = FALSE)
```

```{r}
set.seed(2020)
```

# To do

- More intuitive masking of the arrows based on neighbourhood density.

# Source the function

```{r}
source("plotVectorField.R")
```

# Print the function here, especially note the required packages

```{r}
plotVectorField
```

# Example data using SingleCellExperiment

```{r}
library(scater)
library(scran)
sce = scater::mockSCE(ncells = 5000, ngenes = 100, nspikes = 1)
sce <- logNormCounts(sce)
sce <- runUMAP(sce)
plotUMAP(sce)
```

Assign LFC to be related to the UMAP coordinates, but NA for most values.

```{r}
sce$LFC <-reducedDim(sce, "UMAP")[,1] + 
    abs(reducedDim(sce, "UMAP")[,2]) + 
    rnorm(5000, mean = 0, sd = 0.2)
sce$LFC[sample(ncol(sce), 0.9*ncol(sce))] <- NA

# input coords and outcome
x = reducedDim(sce, "UMAP")[,1]
y = reducedDim(sce, "UMAP")[,2]
val = sce$LFC
```

# Run plotVectorField and view graph

```{r}
res = plotVectorField(x, y, val, mask = FALSE, nbins = 20)
names(res)
lapply(res, class)
res[["plot"]] + ggtitle("Example")
```

And view the arrows dataframe. This is useful for plotting multiple sets of 
arrows on the same plot.

```{r}
head(res[["arrows_df"]])
```

```{r}
sessionInfo()
```