---
title: "An example of incorporating spatial polygon information in SingleCellExperiment"
author: "Shila Ghazanfar"
date: "15/06/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(SingleCellExperiment)
library(scater)
```

# Load functions

```{r}
source("../scripts/segmentation_functions.R")
source("../scripts/spatial_functions.R")
```

# Load some example cell segmentation

```{r}
data = readRDS("../data/dense_segmentation_example.Rds")
boundary_polygons = extractPolygonsFromSegmentation(
    data,
    modes = c("boundaryPolygons"),
    cellFilterFactor = 5,
    minPixels = 50,
    cellPrefix = "cell_",
    cellSuffix = "",
    plot = TRUE,
    verbose = FALSE
)
colnames(boundary_polygons) <- c("x","y","cell","uniqueID")
boundary_polygons <- pruneStraightLines(boundary_polygons, group = "uniqueID")
dim(boundary_polygons)
head(boundary_polygons)
```

# Simulate some expression counts

```{r}
cellnames = unique(as.character(boundary_polygons$uniqueID))
length(cellnames)
genenames = paste0("gene_", 1:10)
exprs = matrix(rpois(length(cellnames)*length(genenames), 50), 
               length(genenames), length(cellnames), 
               dimnames = list(genenames, cellnames))
dim(exprs)
exprs[1:5,1:5]
```

# Build SingleCellExperiment

```{r}
sce = SingleCellExperiment(assays = list(counts = exprs))
colData(sce) <- scater::perCellQCMetrics(sce)
colData(sce)[, "uniqueID"] <- colnames(exprs)
colnames(sce) <- colnames(exprs)
rownames(sce) <- rownames(exprs)
rownames(colData(sce)) <- colnames(sce)
sce
```

# Add cell polygons to colData(sce)

The vertices are assigned as a NumericList, and by splitting the vertices
data.frame by the cell ID "uniqueID".

```{r}
colData(sce)$segmentation_vertices_x = IRanges::NumericList(
    split(
        as.numeric(boundary_polygons$x),
        factor(boundary_polygons$uniqueID, 
               levels = rownames(colData(sce)))
))

colData(sce)$segmentation_vertices_y = IRanges::NumericList(
    split(
        as.numeric(boundary_polygons$y),
        factor(boundary_polygons$uniqueID, 
               levels = rownames(colData(sce)))
))

head(colData(sce))
```

# Given a SingleCellExperiment object, extract vertices for plotting

Using the getSegmentationVerticesDF function.

```{r}
getSegmentationVerticesDF

verticesDF = getSegmentationVerticesDF(colData(sce),
                                       xname = "segmentation_vertices_x",
                                       yname = "segmentation_vertices_y",
                                       othercols = c("uniqueID", "total"))
```

# Plot

Fill the cells by the total counts.

```{r}
g = ggplot(verticesDF, aes(x = segmentation_vertices_x, 
                       y = segmentation_vertices_y,
                       group = uniqueID)) + 
    geom_polygon(aes(fill = total)) + 
    theme_classic() +
    coord_fixed() +
    NULL
g

g + facet_wrap(~ I(total < 500))
```

Fill the cells by the expression for a given gene (not the most well-thought 
out implementation, this could be improved).

```{r}
verticesDF$expression = counts(sce)["gene_1", verticesDF$uniqueID]

g = ggplot(verticesDF, aes(x = segmentation_vertices_x, 
                       y = segmentation_vertices_y,
                       group = uniqueID)) + 
    geom_polygon(aes(fill = expression)) + 
    theme_classic() +
    coord_fixed() +
    NULL
g
```

# To-do

Incorporate an argument to extract from assay slot in getSegmentationVerticesDF

```{r}
sessionInfo()
```